Heuristics and Really Hard Instances for Subgraph Isomorphism Problems

### 2016-01-06 paper/paper.tex
More stuff
More
Graph differently

@@

Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
satisfiable instances tend to be easier but show greater variation than unsatisfiable instances, and
there are exceptionally hard satisfiable instances. (The parallel version of the Glasgow solver is
designed to eliminate these.)

@@

In the top row of \cref{figure:non-induced} we show the phase transition for the non-induced
variant, for patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis)
and target (y-axis) densities. Inside the orange region, at the bottom right of each plot, every
instance is unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the
purple region, at the top left, every instance is satisfiable---we are looking for a sparse pattern
in a dense target (which is easy, since we only have to preserve adjacency, not non-adjacency). The
white band between the regions shows the location of the phase transition: here, roughly half the
instances are satisfiable.

On subsequent rows, we show the average difficulty of different algorithms.  We measure search
nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
solver found the same set of instances difficult.

Satisfiable is easy, until very close to the phase transition. Unsatisfiable is much harder than
satisfiable. This is solver-independent, although VF2 is terrible.

@@

In figure \cref{figure:phase-transition-bands} we show what happens for larger patterns.

\begin{figure}[tb]
    \input{gen-graph-phase-transition-bands.tex}
    \caption{The location of the phase transition, as the pattern grows. Shown are points where
    between 20\% and 80\% of instances are satisfiable. ?? Can we put predictions on here too?}
    \label{figure:phase-transition-bands}
\end{figure}

@@

To predict the location of the phase transition, we can calculate the expected number of solutions
by making a few dubious assumptions involving rounding and independence, and drawing inspiration
from the association graph encoding. Since we are trying to find an \emph{injective} mapping from a
pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are $t^{\underline{p}} = t \cdot (t -
1)\cdots(t - p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the
pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges
being mapped to edges in the target by raising $d_t$ to this power, giving an expected number of
solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}}
\textnormal{,} \] which we plot in ?? somewhere.

@@

@@

>>>>

?? Heuristics

@@

Some stuff about the complexity peak in the Glasgow solver. This looks remarkably similar to random
3SAT problems---compare, for example, Figure 1 of \citet{LeytonBrown:2014}. In particular,
satisfiable instances tend to be easier but show greater variation than unsatisfiable instances, and
there are exceptionally hard satisfiable instances \citep{Smith:1997}.

@@

What if we alter the edge probabilities for both the pattern graph and the target graph?  In the top
row of \cref{figure:non-induced} we show the phase transition for the non-induced variant, for
patterns of order 10, 20 and 30, a target of order 150, and varying pattern (x-axis) and target
(y-axis) edge probabilities. Inside the orange region, at the bottom right of each plot, every
instance is unsatisfiable---here we are trying to find a dense pattern in a sparse target. In the
purple region, at the top left, every instance is satisfiable---we are looking for a sparse pattern
in a dense target (which is easy, since we only have to preserve adjacency, not non-adjacency). The
white band between the regions shows the location of the phase transition: here, roughly half the
instances are satisfiable.

On subsequent rows, we show the average difficulty of different algorithms.  We measure search
nodes. Cannot be used to compare performance of algorithms directly, but this does show that each
solver found the same set of instances difficult.  Satisfiable is easy, until very close to the
phase transition. Unsatisfiable is much harder than satisfiable. This is solver-independent,
although VF2 is terrible.

@@

@@

To predict the location of the phase transition, we can calculate the expected number of solutions
by making a few dubious assumptions involving rounding and independence, and drawing inspiration
from the association graph encoding. Since we are trying to find an \emph{injective} mapping from a
pattern $P = G(p, d_p)$ to a target $T = G(t, d_t)$, there are $t^{\underline{p}} = t \cdot (t -
1) \cdot \ldots \cdot (t - p + 1)$ possible assignments of target vertices to pattern vertices.  We expect the
pattern to have $d_p \cdot \binom{p}{2}$ edges, so we obtain the probability of all of these edges
being mapped to edges in the target by raising $d_t$ to this power, giving an expected number of
solutions of \[ \langle Sol \rangle = t^{\underline{p}} \cdot {d_t}^{d_p \cdot \binom{p}{2}}
\textnormal{,} \] which we plot in the second row of \cref{figure:non-induced}.

We can see that this prediction is generally accurate, except that for very low and very high
pattern densities, we overestimate the satisfiable region slightly. This is because although an
expected number of solutions much below one implies a high likelihood of unsatisfiability, it is not
true that a high expected number of solutions implies that any particular instance is likely to be
satisfiable---a similar behaviour is seen with random constraint satisfaction problems
\citep{Smith:1994,Smith:1996}.

@@

?? Note that it is always best to try to move towards the satisfiable region, even for unsatisfiable
instances. This contradicts the suggestion by \citet{Walsh:1998} that switching heuristics based
upon an estimate of the solubility of the problem may offer good performance.

@@

Dynamic heuristics

Other random models and non-random

Does not give hard instances for graph isomorphism
